Implementing Capon Adaptive Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2012-08-15 HRUIonGPU/jrnl_article.tex
tuffc article

The quadratic complexity with $L$ and linear complexity with $N_K$ can be reduced by realizing that the calculation in (\ref{eq:R}) overlap both across subarrays and in time. The $(i,j)$ element of $\mat{\hat{R}}[n]$ can be calculated from element $(i-1, j-1), \text{for } i,j \in [0, L-1]$  as
\begin{align}
\mat{\breve{R}}_{i,j}[n] &=  \mat{\breve{R}}_{i-1,j-1}[n] - \vec{x}_{i-1}^{\Delta}[n]\vec{x}_{j-1}^{\Delta}[n]^H \nonumber \\
&+ \vec{x}_{i+N_L-1}^{\Delta}[n]\vec{x}_{j+N_L-1}^{\Delta}[n]^H
\end{align}
The same is true in time where
\begin{align}
\mat{\breve{R}}[n] = \mat{\breve{R}}[n-1] - \mat{\breve{R}}[n-K-1] + \mat{\breve{R}}[n-K]
\end{align}
Both sliding across subarrays and in time can be realized simultaneously if we hold a $N_K$-average in time of the outer products $\vec{x}^{\Delta}[n]\vec{x}^{\Delta}[n]^H$ in memory and calculate $\mat{\hat{R}}$ averaged over subarrays from that buffer (ref JOE). However, since $N_K$ typically is small, and we have a limited amount of memory per thread on the GPU, we have chosen to focus on sliding across subarrays only. In that case, the complexity of calculating $\mat{\hat{R}}$ is reduced to $O(L^2)$. Important to notice is that sliding will break independence in the sliding dimension. With our choice of calculating $\mat{\hat{R}}$, calculations are still independent across time, but not across elements. We therefore have selected to use one thread per column in $\mat{\hat{R}}$, and for small $L$s several covariance matrices are calculated per thread block (Figure \ref{fig:gpulayout}). Maximum supported $L$ is with this approach increased to 512, however for small $L$ high $M$ configurations, memory consumption per thread will be to high to achieve full occupancy of the GPU.

@@

Present images before and after Capon weights has been applied. Comment on the effect of selecting different parameters. Present graph showing executing times for Cython-Capon v.s. CUDA-Capon for different choices of parameters. Comment on different choices of parameters change the running times. Temporal averaging for instance, lowers the efficiency since $2*Y_{avg}$ more operations has to be performed per $\mat{\hat{R}}$

@@

(results)?
For a typical cardiac ultrasound image of 80*400 pixels (70 degree sector, 15cm range) acquired using a 2.5MHz, $M=64$ element phased array, the result is 10fps (subarray length $L=M/2$, temporal smoothing over 3 samples). If we do a 2-element pre-beamforming to get the channel count down to 32, the frame rate increases to 44fps. For a 32 element phased array we need less beams to cover the sector (40*400 pixels), hence with the same parameters the frame rate increases to 87fps. The bottleneck of the Capon algorithm is to find the inverse covariance matrix for all pixels. This is an $O(L^3)$ operation, and results show that when L is larger than 16, real time performance is not possible for cardiac imaging with current high-end hardware. The target GPU has been Nvidia’s Quadro 6000, capable of 1Tflops.

@@

% insert where needed to balance the two columns on the last page with
% biographies
\newpage

>>>>

The quadratic complexity with $L$ and linear complexity with $N_K$ can be reduced by realizing that the calculation in (\ref{eq:R}) overlap both across subarrays and in time. Element $(i,j)$ of $\mat{\hat{R}}[n]$ can be calculated from element $(i-1, j-1), \text{for } i,j \in [1, L-1]$ as
\begin{align}
\mat{\breve{R}}_{i,j}[n] &=  \mat{\breve{R}}_{i-1,j-1}[n]  + \sum_{n'=n-K}^{n+K} (x_{i+N_L-1}^{\Delta}[n']x_{j+N_L-1}^{\Delta}[n']^* \nonumber \\
 &- x_{i-1}^{\Delta}[n']x_{j-1}^{\Delta}[n']^H) \label{eq:sliding}
\end{align}
The same is true in time where
\begin{align}
\mat{\breve{R}}[n] &= \mat{\breve{R}}[n-1] + \sum_{l=0}^{N_L-1} (\vec{x}^{\Delta}_l[n+K]\vec{x}^{\Delta}_l[n+K]^H \nonumber \\
&- \vec{x}^{\Delta}[n-K-1]\vec{x}^{\Delta}[n-K-1]^H).
\end{align}
%- \mat{\breve{R}}[n-K-1] + \mat{\breve{R}}[n-K].
Both sliding across subarrays and in time can be realized simultaneously if we hold a $N_K$-average in time of the outer products $\vec{x}^{\Delta}[n]\vec{x}^{\Delta}[n]^H$ in memory and calculate $\mat{\hat{R}}$ averaged over subarrays from this buffer (ref JOE). However, since $N_K$ typically is small, and we have a limited amount of memory per thread on the GPU, we have chosen to focus on sliding across subarrays only (\ref{eq:sliding}). In that case, the complexity of calculating $\mat{\hat{R}}$ is reduced to $O(L^2)$. Important to notice is that sliding will break independence in the sliding dimension. With our choice of calculating $\mat{\hat{R}}$, calculations are still independent across time, but not along diagonals in $\mat{\hat{R}}$. We therefore have selected to use one thread per diagonal in the upper triangle of $\mat{\hat{R}}$, and for small $L$s several covariance matrices are calculated per thread block (Figure \ref{fig:gpulayout}). Maximum supported $L$ is with this approach increased to 512, however for small $L$ high $M$ configurations, memory consumption per thread will be to high to achieve full occupancy of the GPU.

@@

Present images before and after Capon weights has been applied. Comment on the effect of selecting different parameters. Present graph showing executing times for Cython-Capon v.s. CUDA-Capon for different choices of parameters. Comment on different choices of parameters change the running times. Temporal averaging for instance, lowers the efficiency since $2Y_{avg}$ more operations has to be performed per $\mat{\hat{R}}$

@@

(results)?
For a typical cardiac ultrasound image of 80*400 pixels (70 degree sector, 15cm range) acquired using a 2.5MHz, $M=64$ element phased array, the result is 10fps (subarray length $L=M/2$, temporal smoothing over 3 samples). If we do a 2-element pre-beamforming to get the channel count down to 32, the frame rate increases to 44fps. For a 32 element phased array we need less beams to cover the sector ($40 \times 400$ pixels), hence with the same parameters the frame rate increases to 87fps. The bottleneck of the Capon algorithm is to find the inverse covariance matrix for all pixels. This is an $O(L^3)$ operation, and results show that when L is larger than 16, real time performance is not possible for cardiac imaging with current high-end hardware. The target GPU has been Nvidia’s Quadro 6000, capable of 1Tflops.

@@

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage
